### 2016-10-30 paper/ccvpn.tex
Paper directory

\documentclass[conference]{sig-alternate-05-2015}

\usepackage{todonotes,times,amssymb,amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath}
\usepackage{paralist}
\usepackage{multirow,paralist}
\usepackage{fancybox}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{color, colortbl}
\usepackage{transparent}
\usepackage{balance}
\usepackage{url}
\usepackage{graphicx}
\usepackage{paralist}
\usepackage{amssymb, framed, graphics}
\usepackage{bytefield}
\usepackage[group-separator={,}]{siunitx}
\usepackage[underline=true]{pgf-umlsd}
\usepackage{hyphenat}
\usepackage{cite}
\usepackage{fancybox,fancyvrb}
\usepackage{url}
\usepackage{framed}
\usepackage{xspace}
\usepackage{listings}
\usepackage{array}
\usepackage{threeparttable}

\makeatletter
\makeatother

\newcommand{\ccnname}[1]{\path{#1}}
\newcommand{\Adv}{{\sf Adv}}
\newcommand{\mAdv}{\mathsf{Adv}}
\newcommand{\advc}{{\sf Adv}$^C$}
\newcommand{\madvc}{\mathsf{Adv}^C}
\newcommand{\advl}{{\sf Adv}$^L$}
\newcommand{\madvl}{\mathsf{Adv}^L}
\newcommand{\advi}{{\sf Adv}$^I$}
\newcommand{\madvi}{\mathsf{Adv}^I}
\newcommand{\adve}{{\sf Adv}$_E$}
\newcommand{\advo}{{\sf Adv}$_O$}
\newcommand{\advd}{{\sf Adv}$_D$}
\newcommand{\NN}{\bar{N}}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\makeatother
\mathchardef\mhyphen="2D
\newcommand{\E}[1]{\mathrm{E} #1}
\newcommand{\NE}[1]{\mathrm{E}\mhyphen #1}

\newcolumntype{!}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\newcolumntype{N}{@{}m{0pt}@{}}

%\newtheorem{proof}{\textbf{Proof}}
\newtheorem{defn}{\textbf{Definition}}
\newtheorem{define}{\textbf{Definition}}
\newtheorem{thm}{\textbf{Theorem}}
\newtheorem{cor}{\textbf{Corollary}}
\newtheorem{lemma}{\textbf{Lemma}}

\newcolumntype{!}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\newcolumntype{N}{@{}m{0pt}@{}}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\begin{document}

\title{Tunneling in Content Centric Networks}
\numberofauthors{1}
\author{
\alignauthor
%Cesar Ghali \quad Gene Tsudik \quad Christopher A. Wood\titlenote{Supported by the NSF Graduate Research Fellowship DGE-1321846.}\\
%       \affaddr{Computer Science Department, University of California Irvine}\\
%       \email{\{cghali, gene.tsudik, woodc1\}@uci.edu}
}

\maketitle
\pagestyle{plain}

\begin{abstract}
TODO
\end{abstract}

% meat and potatoes
%\input{01-introduction}

%\section{Conclusion and Future Work} \label{sec:conclusion}

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}