### 2016-11-21 paper/ivan/main.tex
writup

\documentclass[conference,letterpaper,10pt]{IEEEtran}
\usepackage{blindtext, graphicx}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\usepackage{microtype}

\usepackage[caption=false,font=footnotesize]{subfig}

\usepackage{fixltx2e}

\usepackage{algorithm2e}

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{CCVPN: Content Centric Virtual Private Networking}

\maketitle

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
%Human Mobility, Group Detection, Group Dynamics, Periodicity, Opportunistic Routing.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\section{Introduction}

\section{Related Work}\label{related}

\begin{itemize}
\item \cite{tsudik2016ac3n}
\item \cite{dibenedetto2011andana}
\item ...
\end{itemize}

\section{CCVPN}\label{metho}

Traditional Virtual Private Networks (VPNs) extend private networks across the Internet. They enable users to send and receive data across shared or public networks as if their computing devices were directly connected to the same private network~\cite{khanvilkar2004virtual}. The goal of CCVPN it to provide its users with the same functionality within the CCN Internet architecture. Therefore, the users can benefit from the features and security of a private network, even though the are not physically under the same private network.

In CCVPN, there are four main entities involved in the virtual private communication: \textit{Consumer}, \textit{Producer}, \textit{Consumer Side Gateway} ($G_c$), and \textit{Producer Side Gateway} ($G_p$). As in usual CCNs, the \textit{Consumer} is the network node that issues an interest for a given content (e.g., file, web page, video) that it wishes to retrieve. The Producer is the network node which originally created such content. $G_c$ and $G_p$ are the edge gateways responsible for ensuring the private communication among distinct domains. As we discuss later on, $G_c$ and $G_p$ can actually be implemented as a single network device, but we present them as separate entities for the purpose of clarity.

\begin{figure*}[!ht]
\centering
\includegraphics[width=2\columnwidth]{ccvpn.pdf}
\caption{CCVPN connectivity architecture}
\label{fig:ccvpn}
\end{figure*}

As depicted in Fig.~\ref{fig:ccvpn}, the devices inside the \textit{Consumer} domain form a physically interconnected private network. Conversely, the devices in the \textit{Producer} domain also form a private network. Therefore, the goal is to create an overlay virtual network that unifies the \textit{Consumer} and the \textit{Producer} domains in way such that the original interests and contents are only visible to the devices inside these two domains. In other words, the interest $I_e$ and the content $C_e$ that are forwarded outside these domains should give no information about the original interest $I_p$ and the actual content $C_p$.

In order to achieve such anonymous communication, $G_c$ is introduced in the \textit{Consumer} domain to encapsulate outgoing interests. Conversely, the $G_p$ is responsible for decapsulating the incoming encapsulated interests and forwarding them in their original form. When the content is forwarded back in response to the interest, $G_p$ encrypts the content before sending it outside the \textit{Producer} domain. Finally, $G_c$ decrypts the received content to its original form and forwards it back towards the \textit{Consumer}. Through the rest of this section we provide a detailed description on how such actions are implemented.

Upon the arrival of a new interest $I_p$, $G_c$ checks its Forwarding Information Base (FIB) to check if the interest prefix is in the list of prefixes for VPN communication. We assume that the FIB must be pre-configured with the list of prefixes which will trigger the VPN communication. Associated with such prefixes are $G_p$'s name and public key ($pk_e$). Therefore, if the prefix of $I_p$ is in VPN prefixes list, $G_P$ runs Algorithm~\ref{alg:interestEncap} to generate a new interest $I_e$ which encapsulates the original interest $I_p$. Firstly, Algorithm~\ref{alg:interestEncap} generates a random symmetric key ($k_r$) which will be used later on to perform the \textit{Content} Encryption/Decryption. Next, it retrieves $G_p$'s name and public key from the FIB. It uses the public key to encrypt the symmetric key $k_r$ and the original interest $I_p$. Then it creates the new interest $I_c$ with $G_p$'s name as the interest name and the generated  ciphertext as payload. Since $I_c$ has  the $G_p$'s name it will be routed towards $G_p$ and since the payload is singed with $G_p$'s public key, only $G_p$ can retrieve the original interest $I_p$ and the symmetric key $k_r$.

\begin{algorithm}[]\label{alg:interestEncap}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\Input{Original interest $I_p$;}
\Output{Encapsulated interest $I_e$;}
%$Ip_{name}$ = getName($I_p$)\\
%storeToPIT($Ip_{name}$)\\
$k_r$ = symmKeyGen($Ip_{name}$);\\
$Gp_{name}$ = retrieveNameFromFIB($I_p$)\\
$pk_e$ = retrievePKFromFIB($I_p$)\\
$payload$ = $Enc_{pk_e}(I_p||k_r)$\\
$I_e$ = createNewInterest($Gp_{name}$, $payload$)\\
storeToPIT($I_e$,$k_r$)\\
\Return $I_e$;\\
\caption{Interest encapsulation (runs on $G_p$)}
\end{algorithm}

After the message is routed towards $G_p$, $G_p$ verifies if the incoming interest name prefix matches its own name and, if it does, $G_p$ runs Algorithm~\ref{alg:interestDecap}, using its own secret key ($sk_e$) to decrypt $I_e$'s payload, which results in the original interest $I_p$ and the symmetric key $k_r$. $G_p$ then stores $I_e$'s name and the symmetric key $k_r$ in its own PIT together with the pending interest $I_p$ and forwards $I_p$. $I_e$'s name and $k_r$ are stored so that they can be used later on to generate the encrypted content $C_e$.

\begin{algorithm}[]\label{alg:interestDecap}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\Input{ Encapsulated interest $I_e$;}
\Input{ Private key $sk_e$;}
\Output{Original interest $I_p$;}
$Ie_{name}$ = getName($I_e$)\\
$cipherText$ = getPayload($I_e$)\\
$I_p||k_r$ = $Dec_{sk_e}(cipherText)$\\
storeToPIT($I_p$,$k_r$,$Ie_{name}$)\\
\Return $I_p$;\\
\caption{Interest decapsulation (runs on $G_c$)}
\end{algorithm}

The original interest $I_p$ is forwarded inside the \textit{Producer} domain until it reaches the \textit{Producer}. The \textit{Producer} responds with the content $C_p$ which is forwarded back to $G_p$.
Upon receiving $C_p$, $G_p$ fetched

\begin{algorithm}[]\label{alg:contentEnc}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\Input{Original content $C_p$;}
\Output{Encrypted content $C_e$;}
$name$ = getName($C_p$)\\
$k_r$ = retrieveKeyFromPIT($name$)\\
$Ie_{name}$ = retrieveNameFromPIT($name$)\\
$payload$ = EncryptThenMAC($k_r$,$C_p$)\\
$C_e$ = createNewContent($Ie_{name}$,$payload$)\\
\Return $C_e$;\\
\caption{Content encryption  (runs on $G_c$)}
\end{algorithm}

\begin{algorithm}[]\label{alg:contentDec}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\Input{Encrypted content $C_e$;}
\Output{Original content $C_p$;}
$Ce_{name}$ = getName($C_e$)\\
$k_r$ = retrieveKeyFromPIT($Ce_{name}$)\\
$cipherText$ = getPayload($C_e$)\\
$C_p$ = Dec($k_r$,$cipherText$)\\
\eIf{$C_p$ == $\perp$} 
    {
    \tcc{MAC verification failed}
    \Return ;\\
    }
    {\Return $C_p$;\\}
\caption{Content decryption (runs on $G_p$)}
\end{algorithm}

For clarity, we have defined a \textit{Consumer} and a \textit{Producer} domain. However, in reality, a single gateway can implement the functions of both $G_c$ and $G_p$. Therefore, consumers and producers can exist in both the domains and interests for contents can be issued from both sides.

\section{Experiments}

\subsection{Experimental Methodology}

\subsection{Results}

\section{Conclusion}\label{conclusion}

\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

\tiny
\bibliographystyle{IEEEtran}
\bibliography{main}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{picture}}]{John Doe}
\blindtext
\end{IEEEbiography}

\end{document}