### 2016-12-29 paper/ccvpn.tex

\documentclass[conference,letterpaper,10pt]{IEEEtran}
\usepackage{blindtext, graphicx, url, paralist}

% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\usepackage{microtype}

\usepackage[tight,footnotesize]{subfigure}
\usepackage{fixltx2e}

\usepackage{multirow}

\usepackage{algorithm2e}

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{CCVPN: Content Centric Virtual Private Networking}

\maketitle

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
%Human Mobility, Group Detection, Group Dynamics, Periodicity, Opportunistic Routing.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\input{01-introduction}
\input{02-overview}
\input{03-related}
\input{04-design}
\input{05-analysis}
\input{06-experiment}
\input{07-security}

\section{Conclusion}\label{conclusion}
TODO

\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

\tiny

\bibliographystyle{IEEEtran}
\bibliography{references}

% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{picture}}]{John Doe}
% \blindtext
% \end{IEEEbiography}

\end{document}