### 2017-01-09 paper/05-sec-analysis.tex

\section{Security Analysis}\label{sec:sec-analysis}
 TODO