### 2017-01-11 paper/05-sec-analysis.tex
better plot ins experiments section

\section{Security Analysis}\label{sec:sec-analysis}

\todo{TODO: Introductory text about the sec analysis}

\begin{definition}
\textit{
An interest encapsulation algorithm $Encapsulate(I_p)$ is an indistinguishable interest encapsulation iff, given any two interests $I_p^1$ and $I_p^2$, chosen by the adversary, and a randomly selected bit $b$, the adversary has only $1/2 + \epsilon$ probability of guessing the value of the bit $b$ when given $I_e^b = Encapsulate(I_p^b)$. Where $\epsilon$ is a negligible factor with regard to the security parameter $k$.
}
\end{definition}

\begin{theorem}
\textit{
Let $Encapsulate_{pk}(I_p)$ denote the interest encapsulation routine described in Algorithm~\ref{alg:interestEncap}. If $Enc_{pk}$, used to construct $Encapsulate_{pk}(I_p)$, is a CPA-secure public-key encryption scheme then $Encapsulate_{pk}(I_p)$ is an indistinguishable interest encapsulation algorithm.
}
\end{theorem}

\textit{\textbf{Proof--}}
\todo{TODO}

\begin{definition}
\textit{
A content encapsulation algorithm $Encapsulate(C_p)$ is an indistinguishable content encapsulation iff, given any two contents $C_p^1$ and $C_p^2$, chosen by the adversary, and a randomly selected bit $b$, the adversary has only $1/2 + \epsilon$ probability of guessing the value of the bit $b$ when given $C_e^b = Encapsulate(C_p^b)$. Where $\epsilon$ is a negligible factor with regard to the security parameter $k$.
}
\end{definition}

\begin{theorem}
\textit{
Let $ContentEnc_{k_r}(C_p)$ denote the content encapsulation routine described in Algorithm~\ref{alg:contentEnc}. If $EncryptThenMAC_{k_r}$, used to construct $ContentEnc_{sk}$, is a CCA-secure symmetric-key encryption scheme, then:
\begin{enumerate}
\item $ContentEnc_{k_r}(C_p)$ is an indistinguishable content encapsulation algorithm;
\item An adversary has only negligible probability of generating a valid fake encapsulated content $I_c'$
\end{enumerate}
}
\end{theorem}

\textbf{\textit{Proof--}}
\todo{TODO}