### 2017-04-06 paper/ccvpn.tex
Updates

\documentclass[conference,letterpaper,10pt]{IEEEtran}
\usepackage{blindtext, graphicx, url, paralist}

% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\usepackage{microtype}

\usepackage{subfigure}
\usepackage{fixltx2e}

\usepackage{multirow}

\usepackage{algorithm2e}
%\usepackage{algorithm,algorithmic}
\usepackage{xcolor}
\newcommand\todo[1]{\textcolor{red}{#1}}

\hyphenation{op-tical net-works semi-conduc-tor}

\newtheorem{definition}{\textbf{Definition}}[section]
\newtheorem{theorem}{\textbf{Claim}}[section]
\newtheorem{corollary}{\textbf{Corollary}}[theorem]
\newtheorem{lemma}[theorem]{\textbf{Lemma}}

\begin{document}

\title{Tunnels to Towers: Secure Virtual Private Networking for CCN}

\maketitle

\begin{abstract}
Content Centric Networking (CCN) is a future Internet architecture which is envisioned as an alternative to the current IP-based model. CCN emphasizes content distribution by making contents directly addressable in a request-based information centric network. An advantage of CCN is that it has some innate privacy friendly features, such as lack of source and destination addresses in packets. However, to be considered a viable future Internet architecture, CCN must offer services for private and anonymous communication that are at least equivalent for  to those present in the IP architecture. Among such, VPNs are a very popular application that enables users to send and receive data across shared or public networks as if their computing devices were directly connected to the same private network. In this work we design, implement and evaluate CCVPN, a content centric virtual private network capable of providing the same functionality of VPNs within the CCN Internet architecture. To the best of our knowledge CCVPN is the first proposal of a VPN alike service for CCNs. In addition to the CCVPN design, we also provide a security analysis and experimental performance evaluation for this new system.
\end{abstract}

\begin{IEEEkeywords}
Content-Centric Networking, VPN, Tunneling
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\input{01-introduction}
\input{02-overview}
\input{03-related}
\input{04-design}
\input{05-sec-analysis}
\input{06-perf-analysis}
\input{07-experiment}

\section{Conclusion}\label{sec:conclusion}
We presented the design, implementation, and assessment of CCVPN, a scheme and
protocol for building secure and private tunnels in CCN. CCPN allows two namespaces
to be bridged by a tunnel across a public network. Unlike point-to-point tunnels,
such as those enabled by secure session protocols, CCVPN allows many consumers
to share the tunnel to access the private namespace. CCVPN is designed with efficiency
in mind: public-key cryptographic operations are kept to a minimum during normal
operation; symmetric-key packet encapsulation algorithms are used to marshall
packets between gateways. Our experimental evidence indicate that CCVPN offers
modest performance in the presence of a variable number of consumers and producers.
The results suggest that CCVPN could be deployed for private namespace tunneling
in real world CCN deployments.

For future work, we plan to implement the CCNxKE protocol from \ref{ccnxke} as a
way to bootstrap symmetric-key tunnels in CCVPN. We will also integrate CCVPN into
real world testbeds to assess its performance for real-world applications. These
include video streaming and large-scale content dissemination applications.

\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

\tiny

\bibliographystyle{IEEEtran}
\bibliography{references}

% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{picture}}]{John Doe}
% \blindtext
% \end{IEEEbiography}

\end{document}